Active Learning from Blackbox to Timed Connectors

### 2016-01-20 paper/paper.tex
.

% personal characters
\newcommand{\rblock}[0]{\circleddash}
\newcommand{\rread}[0]{\rhd}
\newcommand{\rnoread}[0]{\oslash}
\newcommand{\smap}[1]{[{#1}]}
\newcommand{\rempty}[0]{\varnothing}

@@

% --------------------------------------- information -------------------------------------------
\title{Active Learning from Blackbox to Timed Connectors}
\author{
\IEEEauthorblockN{Yi Li\IEEEauthorrefmark{1}, Yiwu Wang\IEEEauthorrefmark{1} and Meng Sun\IEEEauthorrefmark{1}}
\IEEEauthorblockA{
\IEEEauthorrefmark{1}Department of Informatics, School of Mathematical Sciences, Peking University,
Beijing, China\\
liyi\_math@pku.edu.cn, yiwuwang@126.com, summeng@math.pku.edu.cn
}
}

\begin{document}
\maketitle 
\begin{abstract}
  Coordination models and languages play a key role in formally specifying the communication and
  interaction among different components in large-scale distributed and concurrent systems. In this
  paper, we propose an active learning framework to extract timed connector models from black-box
  system implementation. 
  We first introduce parameterized Mealy machine(PMM) as an operational semantic
  model for channel-based coordination language Reo. PMM serves as a bridge
  between Reo connectors and Mealy machines. With the product operator, complex connectors can be
  constructed by joining basic channels and the PMM of connectors can be transformed into Mealy
  machines. Moreover, we adapt L*, a well-known learning algorithm, to timed connectors (in the form
  of Mealy machines). The new algorithm has shown its efficiency in multiple case studies. 
  This framework has been implemented in \texttt{Golang}.
\end{abstract}

@@

Distributed real-time embedded systems (DRES) are reforming our lives with the
Internet of Things(IoT), wherein individual components are composed via \emph{connectors} 
to build systems. Such systems could be distributed logically or physically, which
makes coordination processes even more complicated. In this case, we need to specify the coordination
processes with \emph{coordination languages}, such that formal techniques can be applied to
guarantee their reliability.

@@

In this paper we address this question by proposing a active learning framework as shown in Figure
\ref{fig:howto} to automatically extract timed connector models from blackbox implementations. 
To achieve this goal, we first introduce \emph{parameterized Mealy machine} as a parameterized
semantics for timed Reo connectors, which can be transformed to concrete Mealy machines with a
given alphabet. Then the $L*$ algorithm \cite{DBLP:journals/iandc/Angluin87} is adapted and
optimized to extract Mealy machines with timed action from the connector implementations as
blackboxes.

@@

\begin{itemize}
  \item [-] A \emph{Sync} channel accepts a data item from its source end iff the data item can be
    dispensed through its sink end simultaneously.
  \item [-] A \emph{LossySync} channel is always ready to accept data items. These items will be
    dispensed through its sink end simultaneously if possible, otherwise they will be lost.
  \item [-] A \emph{SyncDrain} channel has two source ends and no sink end. It can accept a data
    item through one of its source end iff a data item is also available to be accepted
    simultaneously through the other end. Then both items will be dropped.
  \item [-] A \emph{FIFO1} channel is an asychronous channel with a buffer cell. It accepts a
    data item whenever the buffer is empty. 
  \item [-] A t-timer channel accepts any data item from its source end, and later dispense it to
    its sink end after a delay of $t$ time units.
    \begin{footnote}
      Here the interpretation of t-Timer is a bit different as in other works like
      \cite{DBLP:conf/sefm/ArbabBBR04, DBLP:conf/tase/Meng12}, but we can easily transform between the
      two kinds of behavior by composing with some other basic channels.
    \end{footnote}
\end{itemize}

Channels are joint together in \emph{nodes}. There are three types of nodes:
\emph{source}, \emph{sink} and \emph{mixed} nodes, depending on whether all channel ends that
coincide on a node are source ends, sink ends or a combination of both.
(see in Figure \ref{fig:node})

@@

\begin{figure}[ht]
  \begin{center}
    \input{./images/connectorexample.tex}
  \end{center}
  \caption{Coordination with Reo Connectors}
  \label{fig:reoconnector}
\end{figure}

@@

\begin{definition}[Mealy machine]
  A Mealy machine is a 6-tuple $(S, s_0, I, O, \delta, \lambda)$ consisting of:
  \begin{itemize}
    \item[-] a finite set of states $S$
    \item[-] an initial state $s_0\in S$
    \item[-] a finite set of inputs $I$
    \item[-] a finite set of outputs $O$
    \item[-] an output function $\delta : S \times  I \rightarrow O$ mapping a pair
      of a state and an input to the corresponding output.
    \item[-] a transition function $\lambda : S \times I \rightarrow S$ mapping a pair of a
      state and an input to the corresponding successor state.
  \end{itemize}
\end{definition}

@@

Besides, we're going to add a ``T'' action in Mealy machines. It indicates that the corresponding
transition will take one time unit to finish.

@@

\begin{definition}[Parameterized Mealy Machine]
  A \emph{Parameterized Mealy Machine} with a parameter $\Sigma$ is defined as a 6-tuple
  $\mathcal{PM}=\langle S, s_0, I, O, \delta, \lambda\rangle$ where 
  \begin{itemize}
    \item[-] The value of $\Sigma$ is a \emph{finite} data set (hereinafter referred as
      alphabet), e.g. $\{a,b\}$
    \item[-] $S$ maps an alphabet to a \emph{finite} set of states.
    \item[-] $s_0$ is the initial state. It satisfies $\forall \Sigma.s_0\in S(\Sigma)$
    \item[-] $I$ is a finite set of source-ends.
    \item[-] $O$ is a finite set of sink-ends.
    \item[-] $\delta$ maps an alphabet to an \emph{output function}. We use
      $\delta(\Sigma):S(\Sigma)\times Input(\Sigma,I,O)\rightarrow Output(\Sigma,O)$ to denote the
      output function.
    \item[-] $\lambda$ maps an alphabet to a \emph{transition function}. We use
      $\lambda(\Sigma):S(\Sigma)\times Input(\Sigma,I,O)\rightarrow S(\Sigma)$ to denote the
      transition function.
  \end{itemize}
\end{definition}

@@

@@

\begin{definition}[Product]
  The product operator \emph{prod} of two PMMs
  \[
  \mathcal{PM'} = prod(\mathcal{PM}_1,\mathcal{PM}_2)\mbox{ if }\mathcal{PM}_2.O\cap
  \mathcal{PM}_1.I=\varnothing
  \]
  is defined as follows.
  \begin{itemize}
  	\item[-] $\forall\Sigma. \mathcal{PM}'.S(\Sigma)=\mathcal{PM}_1.S(\Sigma)\times \mathcal{PM}_2.S(\Sigma)$
    \item[-] $\mathcal{PM}'.I=(\mathcal{PM}_1.I\cup \mathcal{PM}_2.I)\backslash \mathcal{PM}_1.O$
    \item[-] $\mathcal{PM}'.O=(\mathcal{PM}_1.O\cup \mathcal{PM}_2.O)\backslash \mathcal{PM}_2.I$
  \end{itemize}
  \emph{(Here we assume that sink ends of $\mathcal{PM}_1$ can be connected to source ends $\mathcal{PM}_2$, but not
  vise versa)}
  \begin{itemize}
    \item[-] $\mathcal{PM}'.s_0=(\mathcal{PM}_1.s_0, \mathcal{PM}_2.s_0)$
    \item[-] $\forall\Sigma. \mathcal{PM}'.\delta(\Sigma)((s_1,s_2), i)=$
      \begin{displaymath}
        \left\{
        \begin{array}[ht]{ll}
          (Out_1 \cup Out_2)\downarrow_{\mathcal{PM}'.O} & Out_1\neq\rblock\land Out_2\neq\rblock \\
          \rempty & i = T \\
          \rblock & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
      where we have
      \begin{itemize}
        \item[*] $In_1 = i\downarrow_{\mathcal{PM}_1.I}$
        \item[*] $Out_1 = \mathcal{PM}_1.\delta(\Sigma)(s_1,In_1)$
        \item[*] $In_2 = (Out_1 \cup i)\downarrow_{\mathcal{PM}_2.I}$
        \item[*] $Out_2 = \mathcal{PM}_2.\delta(\Sigma)(s_2,In_2)$
      \end{itemize}
    \item[-] $\forall\Sigma. \mathcal{PM}'. \lambda(\Sigma)((s_1,s_2),i)=(s_1',s_2')$
      where we have
      \begin{itemize}
        \item[*] $s_1' = \mathcal{PM}_1.\lambda(\Sigma)(s_1,In_1)$
        \item[*] $s_2' = \mathcal{PM}_2.\lambda(\Sigma)(s_2,In_2)$
      \end{itemize}
  \end{itemize}
\end{definition}

The idea in \emph{production} is quite simple. The output of $\mathcal{PM}_1$ would be provided as part of the
input of $\mathcal{PM}_2$. Time actions T is only executed simultaneously between $\mathcal{PM}_1$ and $\mathcal{PM}_2$.

As mentioned in the notation above, we cannot connect a sink end of $\mathcal{PM}_2$ to a source end of
$\mathcal{PM}_1$, which makes it difficult to construct connectors like alternator in Figure
\ref{fig:reoconnector}. To address the problem, we define a \emph{link} operation to connect sink
ends and source ends in the same connector.

@@

  \begin{itemize}
  	\item[-] $\forall\Sigma. \mathcal{PM}'.S(\Sigma)=\mathcal{PM}.S(\Sigma)$
    \item[-] $\mathcal{PM}'.I=\mathcal{PM}.I\backslash\{\IN\}$
    \item[-] $\mathcal{PM}'.O=\mathcal{PM}.O\backslash\{\OUT\}$
    \item[-] $\mathcal{PM}'.s_0=\mathcal{PM}.s_0$
    \item[-] $\forall\Sigma, i\in
      Input(\Sigma,\mathcal{PM}'.I,\mathcal{PM}'.O).\mathcal{PM}'.\delta(\Sigma)(s,i)=$
      \begin{small}
        \begin{displaymath}
          \left\{
          \begin{array}[h]{lr}
            \rempty & i=T \\
            \mathcal{PM}.\delta(\Sigma)(s,i\cup\{\IN:d\})\downarrow_{\mathcal{PM}'.O} & (*)\\
            \rblock & otherwise \\
          \end{array}
          \right.
        \end{displaymath}
      \end{small}
      where the condition (*) is defined as
      \[
      \exists d.\mathcal{PM}.\delta(\Sigma)(s,i\cup\{\IN:d\})(\OUT)=d
      \]
    \item[-] $\forall\Sigma, i\in
      Input(\Sigma,\mathcal{PM}'.I,\mathcal{PM}'.O).$
      \begin{small}
        \begin{displaymath}
          \mathcal{PM}.\lambda(\Sigma)(s,i)=
          \left\{
          \begin{array}[ht]{lr}
            \mathcal{PM}.\lambda(\Sigma,T) & i=T \\
            \mathcal{PM}.\lambda(\Sigma,i\cup\{\IN:d\}) & (*)\\
            s & otherwise \\
          \end{array}
          \right.
        \end{displaymath}
      \end{small}
  \end{itemize}
\end{definition}

With \emph{link} and \emph{prop} defined, connectors can be constructed by simpler ones in
form of PMMs. It will be transformed into a concrete Mealy machine later, once the alphabet $\Sigma$
is provided.

\section{From Blackbox to Timed Connectors} 
\label{sec:activelearning}
In this section, we use L* algorithm to extract Reo coordinators from blackbox
implementations in 3 steps:
\emph{constructing hypothesis}, \emph{enclosing hypothesis} and \emph{validating hypothesis}. Besides,
here the input and output symbols of the blackbox are provided as $\mathcal{I}$ and $\mathcal{O}$,
and the blackbox implementation is supposed to be equivalent to a Mealy machine where all states are
reachable.

@@

Mealy machines are composed of \emph{states} and \emph{transitions}.
As for blackboxes, input sequence $s\in\mathcal{I}^*$ can be used to denote its state 
after executing $s$. Similarly, for all $a\in\mathcal{I}$, we use $sa$ to denote the successor
state of $s$ after executing $a$.
However, under such notations, there are \emph{infinite} number of states, among which most are
reduplicate.

States are distinguished, usually, by their \emph{subsequent behavior}. Suppose $s_1\neq s_2\in
\mathcal{I}^*$, we say the corresponding of states $s_1$ and $s_2$ are equivalent iff $\forall d\in
\mathcal{I}^+. mq(s_1d)=mq(s_2d)$. Since we can never check it on every $d\in\mathcal{I}^+$, an
alternative way is to use a set of \emph{suffixes} to distinguish them approximately.
\begin{definition}[$D$-Equivalence]
  provided with a set of suffixes $D\subset\mathcal{I}^+$, we say the corresponding states of
  $s_1,s_2\in\mathcal{I}^*$ are $D$-equivalent (denoted as $s_1\sim_D s_2$)
  iff $mq(s_1d) = mq(s_2d)$ for all $d\in D$.
\end{definition}

Now we use \emph{access sequence} to describe the states. Given an input sequence $s$, the access
sequence of $s$ is defined as
\[
acc(s,D)=\min\{s'\in\mathcal{I}^*|s'\sim_D s\}
\]

Formally, an observation table (see in Figure \ref{fig:hypo}) is determined by a tuple $(S,D)$ where
$S\subset\mathcal{I}^*$ is a set of the access sequences of states and $D\subset\mathcal{I}^*$
is a set of suffixes. In such tables, columns are labelled with suffixes in $D$, and rows are
labelled with input sequences. As shown in the example, rows are divided into two parts. The upper
part, labelled with access sequence $s\in S$, denotes the states covered in the current hypothesis.
The lower part, labelled with sequences in $\{sa|s\in S,a\in\mathcal{I}\}\backslash S$ (representing
any potentially unclosed transition) denotes all the transitions outside $s\in S$. A cell with row
label $s$ and column label $d$ is filled with $mq(sd)$.

@@

From a closed observation we can construct a hypothetical Mealy machine in a straight-forward way.
The Mealy machine $\mathcal{H}=\langle S,s_0,\mathcal{I},\mathcal{O},\delta,\lambda\rangle$ is
constructed from $obs$ as follows:
\begin{itemize}
  \item[-] Every state in $S$ corresponds to an input sequence in $obs.S$
  \item[-] $s_0$ corresponds to the empty sequence $\varepsilon$

@@

Algorithm \ref{alg:buildtable} shows how to build an closed observation table.

\begin{algorithm} 
  \caption{CloseTable} 
  \label{alg:buildtable}
  \small
  \KwIn{An oracle interface $mq$, An input actions $\mathcal{I}$, An observation table $obs$}
  \KwOut{A closed observation table $obs$} 
  \Repeat{$unclosed=\varnothing$}
  {
    $next=\{st|\forall s\in obs.S,\forall t\in \mathcal{I}\}\backslash obs.S$\;
    $unclosed=\{seq\in next|\forall a\in obs.S, seq \not\sim_{obs.D} a\}$\;
    $obs.S$ = $obs.S\cup unclosed$\;
  }
  \Return $obs$\; 
\end{algorithm}

Taking a 2-Timer channel as an example, we now illustrate how this algorithm works. 
We assume that the source end of this 2-Timer channel is $A$, the sink end is $B$, and the alphabet
is $\{a\}$. We briefly denote $\{A:a,B:\rnoread\}$ and other input/output actions in form of
$a,\rnoread$.

Firstly, $obs.S$ is initialized with the initial state (denoted by $\varepsilon$), and $obs.D$ is
initialized with all one-step suffixes.
Then we will explore all the access sequences in $obs.S$ and calculate its successors. Here the
$unclosed$ set consists of access sequences that has no equivalent fellow in $obs.S$. 
For example, figure \ref{fig:hypo} shows the $obs$ of a 2-Timer channel after the first
iteration. The five successors of $\varepsilon$ is presented as the bottom part of the table, where
four are equivalent with $\varepsilon$ but $a,\rnoread$ is not. Therefore, we take $a,\rnoread$ as a
brand-new state and all of its successors need further exploration (see in the hypothesis Mealy
machine presented in the following figure).

\begin{figure}[ht]
  \begin{center}
    \begin{displaymath}
      \begin{array}{l||ccccc||c}
        \hline
        & a,\rread & \bot,\rread & a,\rnoread & \bot,\rnoread & T & \mbox{AccessSequence}\\
        \hline\hline
        \varepsilon & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon\\
        \hline
        a,\rread & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon\\
        \bot,\rread & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        \rowcolor[gray]{.9}
        a,\rnoread & \rblock & \rblock & \rblock & \rempty & \rempty & \mbox{NO MATCH}\\
        \bot,\rnoread & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        T & \rblock & \rblock & \rempty & \rempty & \rempty \varepsilon & \varepsilon \\
        \hline
      \end{array}
    \end{displaymath}
    % \input{./images/hypo.tex}
  \end{center}
  \caption{Observation Table and Corresponding Hypothesis}
  \label{fig:hypo}
\end{figure}

In the second iteration, we explore the successors of $a,\rnoread$. Fortunately, now every
successor has a equivalence fellow in $obs.S$ (\emph{itself}). Consequently, the algorithm terminates
with a closed hypothesis where all the unclosed edge shown in Figure \ref{fig:hypo} turn into
self-loop.

>>>>

% personal characters
\newcommand{\rblock}[0]{\circleddash}
\newcommand{\rread}[0]{\rhd}
\newcommand{\rnoread}[0]{\otimes}
\newcommand{\smap}[1]{[{#1}]}
\newcommand{\rempty}[0]{\varnothing}

@@

% --------------------------------------- information -------------------------------------------
\title{Active Learning from Blackbox to Timed Connectors}
\author{
\IEEEauthorblockN{Yi Li, Meng Sun and Yiwu Wang}
\IEEEauthorblockA{
LMAM \& Department of Informatics, School of Mathematical Sciences, Peking University,
Beijing, China\\
liyi\_math@pku.edu.cn, summeng@math.pku.edu.cn, yiwuwang@126.com
}
}

\begin{document}
\maketitle 
\begin{abstract}
  Coordination models and languages play a key role in formally specifying the communication and
  interaction among different components in large-scale distributed and concurrent systems. In this
  paper, we propose an active learning framework to extract timed connector models from black-box
  system implementation. 
  We first introduce parameterized Mealy machine(PMM) as an operational semantic
  model for channel-based coordination language Reo. PMM serves as a bridge
  between Reo connectors and Mealy machines. With the product operator, complex connectors can be
  constructed by joining basic channels and the PMMs of connectors can be transformed into Mealy
  machines. Moreover, we adapt L*, a well-known learning algorithm, to timed connectors (in the form
  of Mealy machines). The new algorithm has shown its efficiency in multiple case studies. 
  This framework has been implemented in \texttt{Golang}.
\end{abstract}

@@

Distributed real-time embedded systems (DRES) are reforming our lives with the
Internet of Things(IoT), wherein individual components are composed via \emph{connectors} 
to build complex systems. Such systems could be distributed logically or physically, which
makes coordination processes even more complicated. In this case, we need to specify the coordination
processes with \emph{coordination languages}, such that formal techniques can be applied to
guarantee their reliability.

@@

In this paper we address this question by proposing an active learning framework as shown in Figure
\ref{fig:howto} to automatically extract timed connector models from blackbox implementations. 
To achieve this goal, we first introduce \emph{parameterized Mealy machine} as a parameterized
semantics for timed Reo connectors, which can be transformed to concrete Mealy machines with a
given alphabet. Then the $L*$ algorithm \cite{DBLP:journals/iandc/Angluin87} is adapted and
optimized to extract Mealy machines with timed action from the connector implementations as
blackboxes.

@@

\begin{itemize}
  \item [-] A \emph{Sync} channel accepts a data item from its source end iff the data item can be
    dispensed through its sink end simultaneously.
  \item [-] A \emph{LossySync} channel is always ready to accept data items. These items will be
    dispensed through its sink end simultaneously if possible, otherwise they will be lost.
  \item [-] A \emph{SyncDrain} channel has two source ends and no sink end. It can accept a data
    item through one of its source end iff a data item is also available to be accepted
    simultaneously through the other end. Then both items will be dropped.
  \item [-] A \emph{FIFO1} channel is an asychronous channel with a buffer cell. It accepts a
    data item whenever the buffer is empty. 
  \item [-] A t-timer channel accepts any data item from its source end, and later dispense it to
    its sink end after a delay of $t$ time units.
    \footnote{
      Here the interpretation of t-Timer is a bit different as in other works like
      \cite{DBLP:conf/sefm/ArbabBBR04, DBLP:conf/tase/Meng12}, but we can easily transform between the
      two kinds of behavior by composing with some other basic channels.
      }
\end{itemize}

Channels are joint together in \emph{nodes}. There are three types of nodes:
\emph{source}, \emph{sink} and \emph{mixed} nodes, depending on whether all channel ends that
coincide on a node are source ends, sink ends or a combination of both.
(see Figure \ref{fig:node})

@@

\begin{figure}[ht]
  \begin{center}
    \input{./images/connectorexample.tex}
  \end{center}
  \caption{An Example of Coordination with Reo Connectors}
  \label{fig:reoconnector}
\end{figure}

@@

\begin{definition}[Mealy machine]
  A Mealy machine is a 6-tuple $(S, s_0, I, O, \delta, \lambda)$ consisting of:
  \begin{itemize}
    \item[-] a finite set of states $S$,
    \item[-] an initial state $s_0\in S$,
    \item[-] a finite set of inputs $I$,
    \item[-] a finite set of outputs $O$,
    \item[-] an output function $\delta :  S \times  I \rightarrow O$ mapping a pair
      of a state and an input to the corresponding output,
    \item[-] a transition function $\lambda : S \times I \rightarrow S$ mapping a pair of a
      state and an input to the corresponding successor state.
  \end{itemize}
\end{definition}

@@

Besides, we add a timed action ``$T$'' in Mealy machines. It indicates that the corresponding
transition will take one time unit to finish.

@@

\begin{definition}[Parameterized Mealy Machine]
  A \emph{Parameterized Mealy Machine} with a parameter $\Sigma$ is defined as a 6-tuple
  $\mathcal{PM}=\langle S, s_0, I, O, \delta, \lambda\rangle$ where 
  \begin{itemize}
    \item[-] The value of $\Sigma$ is a \emph{finite} data set (hereinafter referred as
      alphabet), e.g.  $\{a,b\}$,
    \item[-] $S$ maps an alphabet to a \emph{finite} set of states,
    \item[-] $s_0$ is the initial state. It satisfies $\forall \Sigma.s_0\in S(\Sigma)$,
    \item[-] $I$ is a finite set of source-ends,
    \item[-] $O$ is a finite set of sink-ends,
    \item[-] $\delta$ maps an alphabet to an \emph{output function}. We use
      $\delta(\Sigma):S(\Sigma)\times Input(\Sigma,I,O)\rightarrow Output(\Sigma,O)$ to denote the
      output function,
    \item[-] $\lambda$ maps an alphabet to a \emph{transition function}. We use
      $\lambda(\Sigma):S(\Sigma)\times Input(\Sigma,I,O)\rightarrow S(\Sigma)$ to denote the
      transition function.
  \end{itemize}
\end{definition}

@@

A non-timed input action $a\in Input(\Sigma,I,O)$ can be restricted to $I'\subseteq I$, denoted by
$a\downarrow_{I'}$ which is defined on $Input(\Sigma,I',O)$ and satisfies
\[
\forall i\in I'. a\downarrow_{I'}(i)=a(i)
\]
Restriction on output actions can be defined similarly.

@@

\begin{definition}[Product]
  For two PMMs $\mathcal{PM}_1$ and $\mathcal{PM}_2$, if
  $\mathcal{PM}_2.O\cap\mathcal{PM}_1.I=\varnothing$, their product 
  \[
  \mathcal{PM'} = prod(\mathcal{PM}_1,\mathcal{PM}_2)  \]
  can be defined as follows.
  \begin{itemize}
  	\item[-] $\forall\Sigma. \mathcal{PM}'.S(\Sigma)=\mathcal{PM}_1.S(\Sigma)\times \mathcal{PM}_2.S(\Sigma)$
    \item[-] $\mathcal{PM}'.I=(\mathcal{PM}_1.I\cup \mathcal{PM}_2.I)\backslash \mathcal{PM}_1.O$
    \item[-] $\mathcal{PM}'.O=(\mathcal{PM}_1.O\cup \mathcal{PM}_2.O)\backslash \mathcal{PM}_2.I$
  \end{itemize}
  \emph{(Here we assume that sink ends of $\mathcal{PM}_1$ can be connected to source ends of
  $\mathcal{PM}_2$, but not vise versa)}
  \begin{itemize}
    \item[-] $\mathcal{PM}'.s_0=(\mathcal{PM}_1.s_0, \mathcal{PM}_2.s_0)$
    \item[-] $\forall\Sigma. \mathcal{PM}'.\delta(\Sigma)((s_1,s_2), i)=$
      \begin{displaymath}
        \left\{
        \begin{array}[ht]{ll}
          (\bigcup_{j=1,2} Out_j)\downarrow_{\mathcal{PM}'.O} & i\neq T\land(\bigwedge_{j=1,2} Out_j\neq\rblock) \\
          \rempty & i = T \\
          \rblock & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
      where we have
      \begin{itemize}
        \item[*] $i\in Input(\Sigma,\mathcal{PM}'.I,\mathcal{PM}'.O)$
        \item[*] $In_1 = i\downarrow_{\mathcal{PM}_1.I}$
        \item[*] $In_2 = (Out_1 \cup i)\downarrow_{\mathcal{PM}_2.I}$
        \item[*] $Out_j = \mathcal{PM}_j.\delta(\Sigma)(s_j,In_j)$ for $j=1,2$
      \end{itemize}
    \item[-] $\forall\Sigma. \mathcal{PM}'. \lambda(\Sigma)((s_1,s_2),i)=(s_1',s_2')$
      where for $j=1,2$, $s_j' = \mathcal{PM}_j.\lambda(\Sigma)(s_j,In_j)$.
  \end{itemize}
\end{definition}

The idea in \emph{product} is quite simple. The output of $\mathcal{PM}_1$ would be provided as part of the
input of $\mathcal{PM}_2$. Timed action $T$ is only executed simultaneously between $\mathcal{PM}_1$
and $\mathcal{PM}_2$.

As mentioned above, we cannot connect a sink end of $\mathcal{PM}_2$ to a source end of
$\mathcal{PM}_1$, which makes it difficult to construct connectors like alternator in Figure
\ref{fig:reoconnector}. To solve the problem, we define a \emph{link} operator to connect sink
ends and source ends in the same connector.

@@

  \begin{itemize}
  	\item[-] $\forall\Sigma. \mathcal{PM}'.S(\Sigma)=\mathcal{PM}.S(\Sigma)$
    \item[-] $\mathcal{PM}'.I=\mathcal{PM}.I\backslash\{\IN\}$
    \item[-] $\mathcal{PM}'.O=\mathcal{PM}.O\backslash\{\OUT\}$
    \item[-] $\mathcal{PM}'.s_0=\mathcal{PM}.s_0$
    \item[-] $\forall\Sigma, i\in
      Input(\Sigma,\mathcal{PM}'.I,\mathcal{PM}'.O).\mathcal{PM}'.\delta(\Sigma)(s,i)=$
      \begin{small}
        \begin{displaymath}
          \left\{
          \begin{array}[h]{lr}
            \rempty & i=T \\
            \mathcal{PM}.\delta(\Sigma)(s,i\cup\{\IN\mapsto d\})\downarrow_{\mathcal{PM}'.O} & (*)\\
            \rblock & otherwise \\
          \end{array}
          \right.
        \end{displaymath}
      \end{small}
      where the condition (*) is defined as
      \[
      \exists d.\mathcal{PM}.\delta(\Sigma)(s,i\cup\{\IN\mapsto d\})(\OUT)=d
      \]
    \item[-] $\forall\Sigma, i\in
      Input(\Sigma,\mathcal{PM}'.I,\mathcal{PM}'.O).$
      \begin{small}
        \begin{displaymath}
          \mathcal{PM}.\lambda(\Sigma)(s,i)=
          \left\{
          \begin{array}[ht]{lr}
            \mathcal{PM}.\lambda(\Sigma,T) & i=T \\
            \mathcal{PM}.\lambda(\Sigma,i\cup\{\IN\mapsto d\}) & (*)\\
            s & otherwise \\
          \end{array}
          \right.
        \end{displaymath}
      \end{small}
  \end{itemize}
\end{definition}

With \emph{prod} and \emph{link} defined, connectors can be constructed by composing simpler ones in
the form of PMMs. It can be transformed into a concrete Mealy machine later, once the alphabet
$\Sigma$ is provided.

\section{From Blackbox to Timed Connectors} 
\label{sec:activelearning}
In this section, we use L* algorithm to extract Reo coordinators from blackbox
implementations in 3 steps:
\emph{constructing hypothesis}, \emph{enclosing hypothesis} and \emph{validating hypothesis}. 
The input and output symbols of the blackbox implementation are provided as $\mathcal{I}$ and
$\mathcal{O}$, and the blackbox implementation is supposed to be equivalent to a Mealy machine where
all states are reachable.

@@

Mealy machines consist of \emph{states} and \emph{transitions}.
As for blackboxes, input sequence $s\in\mathcal{I}^*$ can be used to denote its state 
after executing $s$. Similarly, for all $a\in\mathcal{I}$, we use $sa$ to denote the successor
state of $s$ after executing $a$.
However, under such notations, there are \emph{infinite} number of states, among which most are
reduplicate.

States are distinguished, usually, by their \emph{subsequent behavior}. Suppose $s_1\neq s_2\in
\mathcal{I}^*$, we say the corresponding of states $s_1$ and $s_2$ are equivalent iff $\forall d\in
\mathcal{I}^+. mq(s_1d)=mq(s_2d)$. Since we can never check it on every $d\in\mathcal{I}^+$, an
alternative way is to use a set of \emph{suffixes} to distinguish them approximately.
\begin{definition}[$D$-Equivalence]
  Provided with a set of suffixes $D\subset\mathcal{I}^+$, the corresponding states of
  $s_1,s_2\in\mathcal{I}^*$ are $D$-equivalent (denoted as $s_1\sim_D s_2$)
  iff $mq(s_1d) = mq(s_2d)$ for all $d\in D$.
\end{definition}

Now we use \emph{access sequence} to describe the states. Given an input sequence $s$, the access
sequence $acc(s,D)$ of $s$ is defined as the shortest sequence in $\{s'\in\mathcal{I}^*|s'\sim_D
s\}$\footnote{When there are multiple satisfying sequences with the same length, we use a certain tactic
to pick up one from them determinstically.}.

Formally, an observation table (see Figure \ref{fig:hypo} as an example) is determined by a tuple $(S,D)$ where
$S\subset\mathcal{I}^*$ is a set of the access sequences of states and $D\subset\mathcal{I}^*$
is a set of suffixes. In such tables, columns are labelled with suffixes in $D$, and rows are
labelled with input sequences. As shown in Figure \ref{fig:hypo}, rows are divided into two parts. The upper
part, labelled with access sequence $s\in S$, denotes the states covered in the current hypothesis.
The lower part, labelled with sequences in $\{sa|s\in S,a\in\mathcal{I}\}\backslash S$ (representing
any potentially unclosed transition) denotes all the transitions outside $s\in S$. A cell with row
label $s$ and column label $d$ is filled with $mq(sd)$.

@@

From a closed observation table $obs$, we can construct a hypothetical Mealy machine $\mathcal{H}=\langle
S,s_0,\mathcal{I},\mathcal{O},\delta,\lambda\rangle$ in a straight-forward way:
\begin{itemize}
  \item[-] Every state in $S$ corresponds to an input sequence in $obs.S$
  \item[-] $s_0$ corresponds to the empty sequence $\varepsilon$

@@

Algorithm \ref{alg:buildtable} shows how to build a closed observation table.

\begin{algorithm} 
  \caption{EncloseTable} 
  \label{alg:buildtable}
  \small
  \KwIn{An oracle interface $mq$, An input actions $\mathcal{I}$, An observation table $obs$}
  \KwOut{A closed observation table $obs'$} 
  \Repeat{$unclosed=\varnothing$}
  {
    $next=\{st|\forall s\in obs.S,\forall t\in \mathcal{I}\}\backslash obs.S$\;
    $unclosed=\{seq\in next|\forall a\in obs.S, seq \not\sim_{obs.D} a\}$\;
    $obs'.S$ = $obs.S\cup unclosed$\;
    $obs'.D$ = $obs.D$\;
  }
  \Return $obs'$\; 
\end{algorithm}

Taking a 2-Timer channel as an example, we now illustrate how this algorithm works. 
We assume that the source end of this 2-Timer channel is $A$, the sink end is $B$, and the alphabet
is $\{a\}$. We briefly denote $\{A\mapsto a,B\mapsto \rnoread\}$ and other input/output actions in form of
$(a,\rnoread)$.

Firstly, $obs.S$ is initialized with the initial state (denoted by $\varepsilon$), and $obs.D$ is
initialized with all one-step suffixes.
Then we explore all the access sequences in $obs.S$ and calculate its successors. Here
$unclosed$ consists of access sequences that has no equivalent sequence in $obs.S$. 
For example, Figure \ref{fig:hypo} shows the $obs$ of a 2-Timer channel after the first
iteration. The five successors of $\varepsilon$ are presented at the bottom of the table, where
four are equivalent with $\varepsilon$ but $(a,\rnoread)$ is not. Therefore, we take $(a,\rnoread)$ as a
brand-new state and all of its successors need further exploration.

\begin{figure}[ht]
  \begin{center}
    \begin{displaymath}
      \begin{array}{l||ccccc||c}
        \hline
        & (a,\rread) & (\bot,\rread) & (a,\rnoread) & (\bot,\rnoread) & T & acc\\
        \hline\hline
        \varepsilon & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon\\
        \hline
        (a,\rread) & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon\\
        (\bot,\rread) & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        \rowcolor[gray]{.9}
        (a,\rnoread) & \rblock & \rblock & \rblock & \rempty & \rempty & unclosed\\
        (\bot,\rnoread) & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        T & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        \hline
      \end{array}
    \end{displaymath}
    % \input{./images/hypo.tex}
  \end{center}
  \caption{Observation Table and Corresponding Hypothesis}
  \label{fig:hypo}
\end{figure}

In the second iteration, we explore the successors of $(a,\rnoread)$. Fortunately, now every
successor has an equivalence sequence in $obs.S$, \emph{i.e.}, itself. Consequently, the algorithm terminates
with a closed hypothesis where all the unclosed edges turn into
self-loop.
